To be FAIR: Theory Specification Needs an Update

### 2025-02-25 manuscript.Rmd
Update manuscript.Rmd

With regard to the state of Interoperability in psychology, 
Kurt Lewin's -@lewinPsychologyProcessGroup1943 adage "there's nothing as practical as a good theory" paints a hopeful picture of theories as useful tools in psychological researchers' day-to-day work.
But, as we argued, this is not the case.
The examples of X-Interoperability offered in Table \@ref(tab:tabmeehl) illustrate that much can be gained by integrating theory directly into analysis workflows, and by making theory X-Interoperable within software used for analysis.
For example, Interoperable theory could be used
to select control variables for causal inference [@cinelliCrashCourseGood2022],
or to preregister the inferential procedure that would lead to specific modifications of a theory after analyzing empirical data [@peikertReproducibleResearchTutorial2021],
or to derive machine-readable hypotheses [@lakensImprovingTransparencyFalsifiability2021] which could be automatically evaluated through integration testing [@vanlissaUsingEndpointsCheck2023].
Furthermore, theories can be X-Interoperable with each other to enable nesting, or using one theory to clarify elements of another theory.
For example, it should be possible to embed a theory about emotion regulation [e.g., @grossEmotionRegulationCurrent2015] within a theory of emotion regulation development [@morrisRoleFamilyContext2007].

>>>>

With regard to the state of Interoperability in psychology, 
Kurt Lewin's [-@lewinPsychologyProcessGroup1943] adage "there's nothing as practical as a good theory" paints a hopeful picture of theories as useful tools in psychological researchers' day-to-day work.
But, as we argued, this is not the case.
The examples of X-Interoperability offered in Table \@ref(tab:tabmeehl) illustrate that much can be gained by integrating theory directly into analysis workflows, and by making theory X-Interoperable within software used for analysis.
For example, Interoperable theory could be used
to select control variables for causal inference [@cinelliCrashCourseGood2022],
or to preregister the inferential procedure that would lead to specific modifications of a theory after analyzing empirical data [@peikertReproducibleResearchTutorial2021],
or to derive machine-readable hypotheses [@lakensImprovingTransparencyFalsifiability2021] which could be automatically evaluated through integration testing [@vanlissaUsingEndpointsCheck2023].
Furthermore, theories can be X-Interoperable with each other to enable nesting, or using one theory to clarify elements of another theory.
For example, it should be possible to embed a theory about emotion regulation [e.g., @grossEmotionRegulationCurrent2015] within a theory of emotion regulation development [@morrisRoleFamilyContext2007].
